---
title: "p8105_hw2_LC4052"
Author: "Linqi Chu"
output: github_document
date: "2025-09-28"
---

```{r}
library(tidyverse)
library(dplyr)
library(readxl)

```


# Problem 1

## Clean the data in pols-month.csv

```{r}
pols_month = 
  read_csv("./fivethirtyeight_datasets/pols-month.csv",
           na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  separate(mon, c("year", "month","day")) |>
  mutate(
    month = month.name[as.integer(month)],
    president = ifelse(prez_gop == 1, "gop", "dem")
  )|>
  select(-prez_dem, -prez_gop, -day)

```


## Clean the data in snp.csv

```{r}
snp_df = 
  read_csv("./fivethirtyeight_datasets/snp.csv",
           na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  separate(date, c("month", "day","year")) |>
  mutate(
    month = month.name[as.integer(month)],
    year = as.integer(year),
    year = ifelse(year < 50,
                 2000 + year,
                 1900 + year),
    year = as.character(year)) |>
  select(-day) |>
  relocate(year, month, .before = everything()) |> 
  arrange(year, month)


```


## Tidy the unemployment data

```{r}
unemployment_df = 
  read_csv("./fivethirtyeight_datasets/unemployment.csv",
           na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  pivot_longer(
    jan:dec,
    names_to = "month", 
    values_to = "unemployment(%)") |>
  mutate(
    month = factor(month, 
                      levels = c("jan","feb","mar","apr","may","jun",
                                 "jul","aug","sep","oct","nov","dec"),
                      labels = month.name),
    year = as.character(year))

```


## Join the datasets by merging snp into pols, and merging unemployment into the result

```{r}
pols_df = 
  left_join(pols_month, snp_df, by = c("year","month")) |>
  left_join(unemployment_df, by = c("year","month"))

```


## Description of datasets

```{r}
names(pols_month)
names(snp_df)
names(unemployment_df)
names(pols_df)

```

Pols-month contains variables:"year", "month", "gov_gop", "sen_gop", "rep_gop", "gov_dem", "sen_dem", "rep_dem" and "president".
Snp_df contains variables: "year", "month" and "close".
Unemployment_df contains variables: "year", "month" and "unemployment(%)".

The resulting dataset spans from 1947 to 2015 and contains 822 observations, each with 11 features.
It contains variables: "year", "month", "gov_gop", "sen_gop", "rep_gop", "gov_dem", "sen_dem", "rep_dem", "president", "close" and "unemployment(%)".




# Problem 2

## Clean the Mr. Trash Wheel sheet

```{r}
mrtrash_df = 
  read_excel("./trash_collection_data/202409 Trash Wheel Collection Data.xlsx",
             sheet = "Mr. Trash Wheel",
             skip = 1,
             na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  select(-c(15, 16)) |>
  filter(!is.na(dumpster) & !is.na(weight_tons)) |>
  mutate(sports_balls = as.integer(round(sports_balls))) |>
  mutate(trash_wheel = "Mr. Trash Wheel") |>
  mutate(year = as.integer(year))


```


## Clean the Professor Trash Wheel dataset

```{r}
protrash_df = 
  read_excel("./trash_collection_data/202409 Trash Wheel Collection Data.xlsx",
             sheet = "Professor Trash Wheel",
             skip = 1,
             na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster) & !is.na(weight_tons)) |>
  mutate(trash_wheel = "Professor Trash Wheel") |>
  mutate(year = as.integer(year))
  
  
```


## Clean the Gwynnda Trash Wheel dataset

```{r}
gwytrash_df = 
  read_excel("./trash_collection_data/202409 Trash Wheel Collection Data.xlsx",
             sheet = "Gwynnda Trash Wheel",
             skip = 1,
             na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  filter(!is.na(dumpster) & !is.na(weight_tons)) |>
  mutate(trash_wheel = "Gwynnda Trash Wheel") |>
  mutate(year = as.integer(year))


```


# Combine the three dataset

```{r}
trash_data <- bind_rows(mrtrash_df, protrash_df, gwytrash_df)

```


The resulting data has `r nrow(trash_data)` observations, each with `r ncol(trash_data)` variables. The key variables are: `r names(trash_data)`.

The total weight of trash collected by Professor Trash Wheel is `r sum(trash_data[trash_data$trash_wheel == "Professor Trash Wheel", "weight_tons"])`. 
The total number of cigarette butts collected by Gwynnda in June of 2022 is 
`r sum(trash_data[trash_data$trash_wheel == "Gwynnda Trash Wheel", "cigarette_butts"])`.




# Problem 3

## Create a single, well-organized dataset


import, clean, tidy, and otherwise wrangle each of these datasets; check for completeness and correctness across datasets (e.g. by viewing individual datasets and monitoring warning messages); merge to create a single, final dataset; and organize this so that variables and observations are in meaningful orders.

### Clean the zip_codes file

```{r}
zipcodes_df = 
  read_csv("./zillow_data/Zip Codes.csv",
             na = c("NA", ".", "")) |>
  janitor::clean_names()

names(zipcodes_df)

```


### Clean the zip_zori file

```{r}
zori_df = 
  read_csv("./zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv",
             na = c("NA", ".", "")) |>
  pivot_longer(
    cols = 10 : 125,
    names_to = "file_date", 
    values_to = "zori_value") |>
    mutate(file_date = as.Date(file_date)) |>
  rename(county = CountyName) |>                       
  mutate(county = sub(" County$", "", county))  

names(zori_df)


```

```{r}







```

